Map{R}educe in {MPI} for Large-scale Graph Algorithms

### 2010-05-27 paper/results.tex
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@623 3fc257bb-8550-0410-81b3-a5c103da83f7

In Figure~\ref{fig:pr}, we show the performance of the MR-MPI PageRank algorithm
compared to a
distributed-memory matrix-based implementation using the linear
algebra toolkit Trilinos~\cite{Trilinos-Overview} 
and a multi-threaded
implementation
in the Multi-Threaded Graph Library (MTGL)~\cite{MTGL}.
The matrix-based distributed-memory implementation of PageRank
uses Trilinos' Epetra matrix/vector classes to represent the graph and
PageRank vector.
Rows of matrix $A$ and the associated entries of the PageRank vector $x$
are uniquely assigned to processors; a random permutation of the input
matrix achieves processor load balance.
Interprocessor communication gathers $x$ values for matrix-vector
multiplication and sums partial products into the $y$ vector.
Most communication is point-to-point communication,
but some global communication is needed for computing
residuals and norms of $x$ and $y$.
In the multi-threaded MTGL~\cite{MTGL} implementation,
rank propagates via adjacency list traversal
in a compressed sparse-row data structure.
To maintain scalability, code must
be written so that a single thread spawns the loop that processes all
in-neighbors of a given vertex; this detail enables the compiler to generate
hotspot-free code.

Figure~\ref{fig:pr} shows the execution time per PageRank iteration for
R-MAT matrices RMAT-20, RMAT-24 and RMAT-28.  
Converging the PageRank iterations
to tolerance $0.002$ requires five or six iterations.
Several R-MAT matrices of 
each size were generated for the experiments; the average time over the
matrices is reported here.  
The Trilinos implementations show near-perfect
strong scaling for RMAT-20 and RMAT-24.  The MR-MPI 
implementations also demonstrate good strong scaling.  However, MR-MPI's
execution time is at least an order of magnitude greater than the Trilinos
implementation.  This result is due to two factors:  ({\it i}) a 
higher volume of
communication in the MapReduce implementation (where all edges are communicated
in each iteration), compared to the Trilinos implementation (where only
vertex-based data are communicated), and ({\it ii}) out-of-core operations
in MR-MPI were performed because of the page-size restriction, while all
Trilinos operations were performed in-core.  The benefit of MR-MPI's
out-of-core implementation is seen, however, with the RMAT-28 data set, 
which could be solved on smaller processor sets than the Trilinos 
implementation.  For these experiments, the Trilinos implementation required
64 processors for the RMAT-28 data set.

@@

\begin{figure}[h!]
\includegraphics[width=\textwidth]{fig_pagerank.pdf}
\caption{Comparison of PageRank implementations using MapReduce,
Trilinos, and MTGL on R-MAT data sets.}
\label{fig:pr}
\end{figure}

@@

@@

\subsection{Single-source shortest path results}
Discuss SSSP times.

@@

In Figure~\ref{fig:prbig}, we show the performance 
of the various PageRank
implementations on distributed memory and multi-threaded architectures.
The MR-MPI implementation demonstrated good scalability up to 1024 processors; 
however, it required an order-of-magnitude
more execution time than the matrix-based implementations on Redstorm.  
The distributed memory matrix-based
implementations are competitive with the multi-threaded implementation
in MTGL on the Cray XMT.

\begin{figure}[h!]
\includegraphics[width=\textwidth]{fig_pagerank_big.pdf}
\caption{Scalability comparison of PageRank implementations using MapReduce,
Trilinos, and MTGL on R-MAT data sets.}
\label{fig:prbig}
\end{figure}

Similar results were obtained for the Connected Components algorithm, as
we show in Figure~\ref{fig:ccbig}.  As with PageRank, the MR-MPI implementation
showed good scalability up to 1024 processors, but required significantly
more time than the Giant Connected Components algorithms using MTGL and
Trilinos.

\begin{figure}[h!]
\includegraphics[width=\textwidth]{fig_cc_big.pdf}
\caption{Scalability comparison of Connected Components implementations using MapReduce,
Trilinos, and MTGL on R-MAT data sets.}
\label{fig:ccbig}
\end{figure}

>>>>

In Figure~\ref{f:pr}, we show the performance of the MR-MPI PageRank algorithm
compared to a
distributed-memory matrix-based implementation using the linear
algebra toolkit Trilinos~\cite{Trilinos-Overview} 
and a multi-threaded
implementation
in the Multi-Threaded Graph Library (MTGL)~\cite{MTGL}.
The matrix-based distributed-memory implementation of PageRank
uses Trilinos' Epetra matrix/vector classes to represent the graph and
PageRank vector.
Rows of matrix $A$ and the associated entries of the PageRank vector $x$
are uniquely assigned to processors; a random permutation of the input
matrix achieves processor load balance.
Interprocessor communication gathers $x$ values for matrix-vector
multiplication and sums partial products into the $y$ vector.
Most communication is point-to-point communication,
but some global communication is needed for computing
residuals and norms of $x$ and $y$.
In the multi-threaded MTGL~\cite{MTGL} implementation,
rank propagates via adjacency list traversal
in a compressed sparse-row data structure.
To maintain scalability, code must
be written so that a single thread spawns the loop that processes all
in-neighbors of a given vertex; this detail enables the compiler to generate
hotspot-free code.

Figure~\ref{f:pr} shows the execution time per PageRank iteration for
R-MAT matrices RMAT-20, RMAT-24 and RMAT-28.  
Converging the PageRank iterations
to tolerance $0.002$ requires five or six iterations.
Several R-MAT matrices of 
each size were generated for the experiments; the average time over the
matrices is reported here.  
The Trilinos implementations show near-perfect
strong scaling for RMAT-20 and RMAT-24.  The MR-MPI 
implementations also demonstrate good strong scaling.  However, MR-MPI's
execution time is at least an order of magnitude greater than the Trilinos
implementation.  This result is due to two factors:  ({\it i}) a 
higher volume of
communication in the MapReduce implementation (where all edges are communicated
in each iteration), compared to the Trilinos implementation (where only
vertex-based data are communicated), and ({\it ii}) out-of-core operations
in MR-MPI were performed because of the page-size restriction, while all
Trilinos operations were performed in-core.  The benefit of MR-MPI's
out-of-core implementation is seen, however, with the RMAT-28 data set, 
which could be solved on smaller processor sets than the Trilinos 
implementation.  For these experiments, the Trilinos implementation required
64 processors for the RMAT-28 data set.

@@

\begin{figure}[h!]
\includegraphics[width=\textwidth]{fig_pagerank.pdf}
\caption{Comparison of PageRank implementations using MapReduce,
Trilinos, and MTGL on R-MAT data sets.}
\label{f:pr}
\end{figure}

@@

KDD NOTE:  Need to add XMT results.

\subsection{Connected Components}
Discuss Luby maximally independent sets times.

KDD NOTE:  Need to add XMT results.

@@

\subsection{Single-source shortest path results}

The execution times for the single-source shortest path algorithms
(Figures~\ref{fig:sssp} and \ref{fig:sssp2})
are shown in Figure~\ref{f:sssp}.  The results show the benefit of
using the enhanced algorithm, providing at least a 17\% (and often greater)
reduction in execution time due to reduced communication; only the updated
distances are communicated throughout most of the enhanced algorithm.
However, the execution times are still large compared to multi-threaded
implementations; for example, Madduri et al.~\cite{Madduri07} report
execution times of only 11 seconds on 40 
processors for a multithreaded implementation on a Cray MTA-2
(the predecessor of the XMT) on graphs with one billion edges.

\begin{figure}[h!]
\includegraphics[width=\textwidth]{fig_sssp.pdf}
\caption{Execution times for SSSP using MR-MPI with R-MAT matrices.  
Both the original algorithm (Figure~\ref{fig:sssp}) and the enhanced 
algorithm (Figure~\ref{fig:sssp2}) are included.}
\label{f:sssp}
\end{figure}

To compare our MR-MPI implementation with a wider set of implementations,
we performed experiments MR-MPI, Hadoop, PBGL (Parallel
Boost Graph Library)~\cite{PBGL} and a multi-threaded implementation on the
Cray XMT using two web graphs: {WebGraphA} with 13.2M
vertices and 31.9M edges, and {WebGraphB} with 187.6M vertices and 531.9M edges.
In Figure~\ref{f:ssspA}, we show execution times for the SSSP algorithm
using {WebGraphA}.   Like our MR-MPI implementation, the Hadoop 
implementation is a Bellman-Ford-style~\cite{Bellman58,Ford62}  algorithm.  
The XMT and PBGL implementations are based on 
delta-stepping~\cite{MeyerSanders98}, and do not require full iterations
over the entire edge list to advance a breadth-first search.  
We observe that the MR-MPI
implementation runs in less time than the Hadoop implementation, but requires
significantly more time than the XMT and PBGL implementations.  
The Bellman-Ford-style iterations are especially harmful to the MR-MPI and
Hadoop implementations for {WebGraphB}, which required 109 iterations to 
complete the search; 
execution times for this data set are show in Table~\ref{t:ssspB}.

\begin{figure}[h!]
\includegraphics[width=\textwidth]{fig_ssspA.pdf}
\caption{SSSP using MR-MPI for WebGraphA with
13.2M vertices and 31.9M edges.  Runtimes using Hadoop and PBGL
are also shown.}
\label{f:ssspA}
\end{figure}

\begin{table}
\begin{tabular}{|l|c|r|}
\hline
Implementation & Number of Processes & SSSP Execution Time \\
\hline
Hadoop & 48  & 38,925 secs.\\
MR-MPI & 48 &  13,724 secs.\\
XMT/C  & 32 &  37 secs.\\
%Out-of-core MR-MPI & 96 &   8,358 secs.\\
%Out-of-core MR-MPI & 64 &  12,882 secs.\\
%Out-of-core MR-MPI & 100 &  6,280 secs.\\
\hline
\end{tabular}
\caption{Execution times for SSSP with {WebGraphB}.}
\label{t:ssspB}
\end{table}

@@

In Figure~\ref{f:prbig}, we show the performance 
of the various PageRank
implementations on distributed memory and multi-threaded architectures.
The MR-MPI implementation demonstrated good scalability up to 1024 processors; 
however, it required an order-of-magnitude
more execution time than the matrix-based implementations on Redstorm.  
The distributed memory matrix-based
implementations are competitive with the multi-threaded implementation
in MTGL on the Cray XMT.

\begin{figure}[h!]
\includegraphics[width=\textwidth]{fig_pagerank_big.pdf}
\caption{Scalability comparison of PageRank implementations using MapReduce,
Trilinos, and MTGL on R-MAT data sets.}
\label{f:prbig}
\end{figure}

Similar results were obtained for the Connected Components algorithm, as
we show in Figure~\ref{f:ccbig}.  As with PageRank, the MR-MPI implementation
showed good scalability up to 1024 processors, but required significantly
more time than the Giant Connected Components algorithms using MTGL and
Trilinos.

\begin{figure}[h!]
\includegraphics[width=\textwidth]{fig_cc_big.pdf}
\caption{Scalability comparison of Connected Components implementations using MapReduce,
Trilinos, and MTGL on R-MAT data sets.}
\label{f:ccbig}
\end{figure}
